---
title: "Lab 1"
author: "Your  Name"
output: html_notebook
---

This lab covers:

Calculator & Objects

* running basic calculations
* assigning and returning variables
* calculate using variables
* re-assign variables
* determine the type of an object

Numeric Vectors

* create a vector with specific numbers
* create a vector with a range of numbers
* basic math on vectors
* combining vectors -- recycling

Logical Vectors

* Logical operators
* math on logical 

Subsetting
For loops
If statements


We can ask for input from a user in R by using the funciton "readline." For example, the code below creates a custom function called "ask" that asks for a name (saved as the variable `z`) then prints "Hello <name>!"
```{r}
ask <- function(){
  z <- readline("Enter your name: ")
  print(paste("Hello ", z, "!", sep=""))
}

ask()
```


1. Ask for a positive number and return a vector of all the integers between 1 and the input. For example, if the user picks 5, your code should return `1 2 3 4 5`
```{r}
ask <- function(){
  num <- readline("Enter a number: ")
  num <- as.numeric(num) # Note that we're telling R to interpret the input (num) as a number!
  return (c(1:num))
}
ask()
```

2. Ask for a number and return a vector of all the even numbers between 1 and the input. Hint: You can do this with a loop, or you can try reading about the function `seq()` by typing `?seq()` into the console.
```{r}
ask <- function(){
  num <- readline("Enter a number: ")
  num <- as.numeric(num)
  ans <- seq(2, num, by=2)
  return(ans)
}
ask()
```

3. Ask for a positive integer n. Return the sum: 1 + 1∕2 + 1∕3 + 1∕4 + ⋯ + 1∕n. Hint: learn about the `sum()` function.
```{r}
ask <- function(){
  num <- readline("Enter a number: ")
  num <- as.numeric(num)
  ans <- sum(1/(1:num))
  return(ans)
}
ask()
```
4. Ask for a positive integer n. Return a vector with elements: 1, 1/2, 1/3, 1/4, ... , 1/n.
```{r}
ask <- function(){
  num <- readline("Enter a number: ")
  num <- as.numeric(num)
  ans <- c(1/1:num)
  return(ans)
}
ask()
```

5. Using the vectors provided, ask for a number and print the name of the person at that index, and whether or not they like cake.
```{r}
people <- c("Alice", "Bob", "Charlie", "David", "Eve")
likesCake <-  c(TRUE, FALSE, TRUE, TRUE, FALSE)

ask <- function(){
  num <- readline("Enter a number between 1 and 5: ")
  num <- as.numeric(num)
  
  name <- people[num]
  
  if(likesCake[num]){
    print(paste(name, "likes cake"))
    
  } else {
    print(paste(name, "does not like cake"))
  }
}
ask()
```

6. Ask for a start and end, then return all elements of the provided vector between those indices.
```{r}
vector <- c("red", "orange", "yellow", "green", "blue", "purple")
ask <- function(){
  start <- readline("Enter start: ")
  end <- readline("Enter end: ")
  ans <- vector[start:end]
  return(ans) 
}
ask()
```

7. Ask the user for 5 numbers and add them one by one to a vector.
```{r}
ask <- function(){
  vector <-  c()
  for(i in 1:5){
    num <- readline("Enter a number: ")
    num <- as.numeric(num)
    vector[i] <- num
  }
  return(vector)
}
ask()
```

8. Print a vector that is `TRUE` when the sum of the given vectors is greater than 5 and `FALSE` otherwise.

```{r}
vec1 <- c(0,1,2,4,1,5,6,2,3)
vec2 <- c(4,3,5,4,8,9,4,2,3)

sumFive <- vec1+vec2 < 5

print(sumFive)
```

9. Ask for the total of a bill and return the amount of a 15% tip. The `round` function is useful for cleaning up the result to an even penny. Use `?round` to see the help documentation.
```{r}
ask <- function(){
  num <- readline("Enter total: $")
  num <- as.numeric(num)
  ans <- round(num*0.15, 2)
  return(ans)
}
ask()
```

10. Ask for a user’s birth year and print the age they will turn this year. You can write the current year directly without trying to determine it externally (it is possible to determine the current year, but was not covered in Chapter 2).

```{r}
ask <- function(){
  num <- readline("Enter your birth year: ")
  num <- as.numeric(num)
  ans <- 2020-num
  print(paste("You turned (or are turning)", ans, "years old this year."))
}
ask()
```

11. Write a function which asks for a number and determines if it is a whole number. Print a message displaying the result using `print`.
```{r}
ask <- function(){
  num <- readline("Enter a number: ")
  num <- as.numeric(num)
  if (num == round(num)) print("Whole number")  
  if (num != round(num)) print("Not a whole number")
}
ask()
```

12. The factorial of an integer is the product of all the positive integers less than or equal to it. For example, the factorial of 4 is equal to 4 ∗ 3 ∗ 2 ∗ 1=24. There is a function factorial in base R for calculating these. Ask for a number and return the factorial, without using the R function. Hint: The function `prod` may be helpful.
```{r}
ask <- function(){
  num <-  as.numeric(readline("Enter a number: "))
  return(prod(num:1))
}
ask()
```

13. Let's play a game! Randomly pick a number between 1 and 10, then give the user 3 attempts to guess correctly. Your game should tell the user when they've won, if they guessed wrong and should try again, or if they've lost. Hint: you'll need the function `sample`. Use `?sample` to learn about it! You'll also want to use `break` to stop your loop when they win.
```{r}
play <- function(){
  secret <- sample(1:10, 1)
  
  for (try in 1:3){
    guess <- readline("Guess my number: ")
    guess <- as.numeric(guess)
    if (guess == secret){
      print("You win!")
      break
    }
    else if (try < 3){
      print("Nope! Try again...")
    }
    else {
      print("Sorry! I was thinking of:")
      print(secret)
    }
  }
}

play()
```

BONUS CHALLENGE. Re-write your game to pick a number between 1 and 20, give them 5 guesses, and give hints like "warmer" or "colder" after each guess.
```{r}
play <- function(){
  secret <- sample(1:20, 1)
  num_guesses = 5
  
  last_guess = 0 # track the previous guess
  
  for (try in 1:num_guesses){
    guess <- readline("Guess my number: ")
    guess <- as.numeric(guess)
    if (guess == secret){
      print("You win!")
      break
    }
    else if (try < num_guesses && try > 1){
      if (abs(last_guess-secret) < abs(guess-secret)){ # if they got farther away
        print("Colder...")
      } else if (abs(last_guess-secret) > abs(guess-secret)){ #if they got closer
        print("Warmer...")
      }
      
      last_guess = guess
    }
    else if (try < num_guesses){ # if they guess the same distance OR it's the first guess
      print("Nope! Try again...")
    }
    else {
      print("Sorry! I was thinking of:")
      print(secret)
    }
  }
}

play()
```